---
title: "neurogenes synteny project"
csl: plos.csl
output:
  pdf_document: default
  github_document: default
  html_document: default
bibliography: Bib.bib
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	echo=FALSE,
	dpi=300,
	cache=TRUE
	)
```

```{r preliminaries}
#library( tidyverse )
```


1. 
#Paper notes
###Moroz et al., 2014  [@Moroz2014]
Table 34S
-structure of figure shows that it is from bilaterian perspective (closer to bilateria more rectangles filled in)  
-remember there are also random absences in cnidaria; the mirror of bilateria, but all the proteins are characterized from bilateria  
-no examples where ctenophores or sponges don't have something present in fungi, capsaspora, monosiga  
-Suppl Table 12as - it's possible didn't use Amphimedon genome, but Amphimedon was covered by the other papers and I crossreferenced them.  

Mentioned in text:  
Not in ctenophores:   
neurogenin  
NeuroD  
Achaete-scute  
REST  
HOX  
Otx  

-not that much overlap in genes looked at by Riesgo vs Moroz  

###Ryan et al., 2013  [@Ryan2013]
*netrin, slit, unc-5 (axon guidance) not in Mnemiopsis or Amphimedon*  
-used genomes, since based on Alie and Manuel 2010  
Supplementary Table S17: Presence and absence of post-synaptic genes  - pretty much Alie and Manuel 2010
Supplementary Table S19: Presence and absence of Dopamine / Norepinephrine /Epinephrine Biosynthetic Pathway components   

-are seqs of the animals in S17 genomes? Unsure, but all animals in table have genomes (and the Mle seqs are from the genome)  
-AMPA iGluR and NMDA iGluR included as iGluR  

###Alie and Manuel, 2010  [@Alie2010]
-used genomes  
-Ryan built on Fig. 1. Cross ref with current data to make sure have everything.  
-Only use Monosiga, Trichoplax, Amphimedon, Nematostella, Hydra, Homo  
  Capitella (3 absences), Drosophila (2 absences), Homo very similar with few differences  
  Unicellular animals mostly missing everything (except B-cat and PMCA). Start with Monosiga which has more things.
B-cat and PMCA are ancient - interesting?  
AMPAR and NMDAR collapsed into iGluR in table; presence of one of these trumped absence of the other  
PKC alpha-beta-gamma = PKC on table  

#[Table](https://github.com/jasminelmah/compgen/blob/master/neurogenes_table.numbers)  
_Table abbreviations_  
DBH - dopamine-B-hydroxylase  
DDC - DOPA decarboxylase  
TH - tyrosine hydroxylase  
TPH - tryptophan hydroxylase  
PAH - phenylalanine hydroxylase  
GAD - glutamate decarboxylase  
Qdpr -  quinoid dihydropteridine reductase,  
Slc18A2 = Homo sapiens solute carrier family 18 member 2,  
Pnmt = phenylethanolamine N-methyltransferase  

Missing domains  
Piccolo - Pleurobrachia - missing ZF (Moroz et al., 2014)  
Erbin - Pleurobrachia - missing PDZ (Moroz et al., 2014)  

Species names written to the broadest level - eg. Monosiga brevicollis in Riesgo et al but only Monosiga in Moroz, so put Monosiga only  

Many entries have NA but if combine:  
Salpingoeca + Monosiga = Choanoflagellida  
Pleurobrachia + Mnemiopsis = Ctenophora  
Amphimedon + Oscarella = Porifera  
Nematostella + Hydra = Cnidaria  
Get only 4 entries that have an NA.  
*(What about 0/1s (conflicting info?)*  >> decided to transform 0/1s into NA

Loss_Status:
P1C0: present in Porifera, absent in Ctenphora - 6 instances  
C1P0: present in Ctenophora, absent in Porifera - 3 instances  
T0: absent in Trichoplax but present in Ctenophora or Porifera - 5 instances  

There is only 2 instances where Capsaspora has a 1 while choanoflagellates have 0: GABAR and DDC. Don't use column in second iteration  
'Fungi' is very vaguely defined - don't use column in second iteration  
Stopped at Delta catenin

Make new table where all 0/1s or missing_domains (i.e. not 0,1,NA) into NA  

Create a new table where species for Ctenophora, Porifera combined: https://stackoverflow.com/questions/14563531/combine-column-to-remove-nas  

#Papers+Links
Riesgo et al., 2014  
https://academic.oup.com/mbe/article/31/5/1102/993377  
Neural genes Fig  
https://academic.oup.com/view-large/figure/74385341/msu057f3p.jpeg  

Moroz et al., 2014  
https://www.nature.com/articles/nature13400  
Table 34S:  
https://media.nature.com/original/nature-assets/nature/journal/v510/n7503/extref/nature13400-s1.pdf  

Ryan et al, 2013  
http://science.sciencemag.org/content/342/6164/1242592  
Suppl Mat:  
http://science.sciencemag.org/content/sci/suppl/2013/12/11/342.6164.1242592.DC1/Ryan.SM.pdf  

Alie and Manuel, 2010  
https://bmcevolbiol.biomedcentral.com/articles/10.1186/1471-2148-10-34  

Srivastava et al., 2010  
Suppl.S8.9
https://media.nature.com/original/nature-assets/nature/journal/v466/n7307/extref/nature09201-s1.pdf


2. Should we be expecting these genes in these animals?  
Why should they use similar genes?  
How misguided is this approach? What is the true question implied by this approach?  


##References  
