---
title: "neurogenes synteny project"
csl: plos.csl
output:
  pdf_document: default
  github_document: default
  html_document: default
bibliography: Bib.bib
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	echo=FALSE,
	dpi=300,
	cache=TRUE
	)
```

```{r preliminaries}
#library( tidyverse )
```


1. 
#Paper notes - neural genes presence/absences  

###Moroz et al., 2014 [@Moroz2014]  
-Table 34S:
-structure of figure shows that it is from bilaterian perspective (closer to bilateria more rectangles filled in)  
-remember there are also random absences in cnidaria; the mirror of bilateria, but all the proteins are characterized from bilateria  
-no examples where ctenophores or sponges don't have something present in fungi, capsaspora, monosiga  
-Suppl Table 12as - it's possible didn't use Amphimedon genome, but Amphimedon was covered by the other papers and I crossreferenced them.  

Mentioned in text:  
Not in ctenophores:   
neurogenin  
NeuroD  
Achaete-scute  
REST  
HOX  
Otx  

-not that much overlap in genes looked at by Riesgo vs Moroz  

###Ryan et al., 2013 [@Ryan2013]  
*netrin, slit, unc-5 (axon guidance) not in Mnemiopsis or Amphimedon*  
-used genomes, since based on Alie and Manuel 2010  
Supplementary Table S17: Presence and absence of post-synaptic genes  - pretty much Alie and Manuel 2010
Supplementary Table S19: Presence and absence of Dopamine / Norepinephrine /Epinephrine Biosynthetic Pathway components   

-are seqs of the animals in S17 genomes? Unsure, but all animals in table have genomes (and the Mle seqs are from the genome)  
-AMPA iGluR and NMDA iGluR included as iGluR  

###Alie and Manuel, 2010 [@Alie2010]  
-used genomes  
-Ryan built on Fig. 1. Cross ref with current data to make sure have everything.  
-Only use Monosiga, Trichoplax, Amphimedon, Nematostella, Hydra, Homo  
  Capitella (3 absences), Drosophila (2 absences), Homo very similar with few differences  
  Unicellular animals mostly missing everything (except B-cat and PMCA). Start with Monosiga which has more things.
B-cat and PMCA are ancient - interesting?  
AMPAR and NMDAR collapsed into iGluR in table; presence of one of these trumped absence of the other  
PKC alpha-beta-gamma = PKC on table  

###Srivastava et al, 2010 [@Srivastava2010]  
-has extensive tables of presence/absence neural genes (neurogenesis S.8.9.1, presynaptic S.8.9.2, postsynaptic S8.9.3, neurosecretory S. 8.9.4 and see Fig. 2.8.9.1 for "synapse diagram"). BUT classifies by Holozoa, Metazoa, Eumetazoa, Bilateria. Can figure out in sponges if Metazoa (?) but too vague? Need specific genomes?   

#[Table](https://github.com/jasminelmah/compgen/blob/master/neurogenes_table.numbers)  
_Table abbreviations_  
DBH - dopamine-B-hydroxylase  
DDC - DOPA decarboxylase  
TH - tyrosine hydroxylase  
TPH - tryptophan hydroxylase  
PAH - phenylalanine hydroxylase  
GAD - glutamate decarboxylase  
Qdpr -  quinoid dihydropteridine reductase,  
Slc18A2 = Homo sapiens solute carrier family 18 member 2,  
Pnmt = phenylethanolamine N-methyltransferase  

Missing domains  
Piccolo - Pleurobrachia - missing ZF (Moroz et al., 2014)  
Erbin - Pleurobrachia - missing PDZ (Moroz et al., 2014)  

Species names written to the broadest level - eg. Monosiga brevicollis in Riesgo et al but only Monosiga in Moroz, so put Monosiga only  

Many entries have NA but if combine:  
Salpingoeca + Monosiga = Choanoflagellida  
Pleurobrachia + Mnemiopsis = Ctenophora  
Amphimedon + Oscarella = Porifera  
Nematostella + Hydra = Cnidaria  
Get only 4 entries that have an NA.  
(What about 0/1s (conflicting info?)  >> decided to transform 0/1s into NA

Loss_Status:
P1C0: present in Porifera, absent in Ctenphora - 6 instances  
C1P0: present in Ctenophora, absent in Porifera - 3 instances  
T0: absent in Trichoplax but present in Ctenophora or Porifera - 5 instances  

Second Iteration:  
-There is only 2 instances where Capsaspora has a 1 while choanoflagellates have 0: GABAR and DDC. Don't use column in second iteration  
-'Fungi' is very vaguely defined - don't use column in second iteration  
-Stopped at Delta catenin
-Make new table where all 0/1s or missing_domains (i.e. not 0,1,NA) into NA  

Via R
Create a new table where species for Ctenophora, Porifera combined: https://stackoverflow.com/questions/14563531/combine-column-to-remove-nas  

#Synteny programs/papers  
ghost locus hypothesis:  
Ramos et al., 2012 [@Ramos2012]  
https://www.sciencedirect.com/science/article/pii/S0960982212009888  
-first ghost locus paper - parahox, Amphimedon + Trichoplax  

Fortunato et al, 2014 [@Fortunato2014] - sycon Parahox  
https://www.nature.com/articles/nature13881  
-second ghost locus paper?  

Ferrier 2015 [@Ferrier2015] (review)  
https://academic.oup.com/bfg/article/15/5/333/1741867  


Reviews:  
Liu et al 2018 [@Liu2018]  
https://bmcbioinformatics.biomedcentral.com/articles/10.1186/s12859-018-2026-4  
 
Liu Results:  
-Pos control (default params): comparison of genome with itself: SynChro (most synteny), DAGchainer/i-ADHoRe, MCScanX 
-Fragmentation: break 2 species genomes [C. elegans, C. briggsae, S. rattis, S. stercoralis] (diff gene density) into diff fragment lengths. pos corr between error rate and level of fragmentation except for Satsuma. For anchor-based programs, fragmentation has biggest effect on MCScanX and lowest on SynChro.  
-Find synteny between sister species: Satsuma worst (diff with alignments/nt homology), Anchor-based programs found much higher synteny - can't access Add. file 2 Table S1 that describes this.  
-Fragmented species and compared to its own genome: MCScanX worst. The rest of anchor-based progs performed similarly.k SyncChro often better or second best.Degree of error differed per species.  
-high gene density, less error (more anchors)  
-fragmented assemblies lead to erroneous GO terms.  
-reference-guided assembly methods rely on assumption of synteny = synteny errors.  

Bottom line:  
-most synteny programs designed with assumption working on complete high quality genomes. 
-SynChro best for fragmented assemblies in general, imo.  
-Satsuma does not have a positive correlation with error rate and level of fragmentation, but performed poorly comparing two closely related species due to failure of alignment - go with anchor-based.   
-need N50 of at least 200 kb and gene density 290 genes/Mb, or N50 1 Mb and gene density 200 genes/Mb for error rate < 5%. Higher N50 for genomes with less gene density or many paralogues/expansion of gene families.   
-annotation qualityt really doesn't matter just need high genome assembly contiguity.  
-more paralogues less synteny finding?? may have misunderstood.  

##To do:
-Check to see whether genomes assembled via ref to related species -> leads to false synteny due to assumption of shared synteny  
-Check N50s of my genomes  

**Scaffold N50s of my genomes**  
Mnemiopsis (from orig @Ryan2013): 187 kb  
Pleurobachia (Suppl Table 5S from @Moroz2014): 20.607 kb  
Amphimedon orig, [Ensembl (https://metazoa.ensembl.org/Amphimedon_queenslandica/Info/Annotation/): 120 kb >> likely improved, looking for Aqu2.1 (or is this only a re-annotation not reassembly?)  
  - Slightly confused with scaffold N50 stats in @Srivastava2010 suppl Table S2.3.2  
Oscarella carmela (Suppl from @Nichols2012): 5.897 kb!! :O  

#SYNTENY ANALYSIS PLAN  
Test programs against the same genome - concentrate on the four cteno and por genomes (?).  
Use a normal synteny program to identify genes surrounding gene of interset.  
Then follow Ramos procedure.  



#First run attempt with Synchro  
-Aqu with Aqu: A vs B    
.gbff from Genbank = ".dat" for Synchro   
f.dat lat file from embl = ".embl" for Synchro    
Found .dat aka .embl [here] (https://www.ebi.ac.uk/ena/data/view/GCA_000090795.1)  
Found .gbff aka genbank .dat [here] (ftp://ftp.ncbi.nlm.nih.gov/genomes/all/GCA/000/090/795/GCA_000090795.1_v1.0)    

genbank: it's possible .dat is outdated. .gbff has replaced .gbk and Synchro. Yeah, C. elegans gbff doesn't work. 
Archived [genbank] (ftp://ftp.ncbi.nlm.nih.gov/genomes/archive/old_genbank) and [refseq files] (ftp://ftp.ncbi.nlm.nih.gov/genomes/archive/old_refseq/) scaffolds only, no annotation files - before annotation done?  

**Big wall: Can't get file types to work.**  

#First run attempt with DAGchainer (http://dagchainer.sourceforge.net/)  
Popular. MCScan is popular too but performed worse.    
**make is not working. Get error:**    
``  
g++ -o dagchainer. dagchainer.cpp -Wno-deprecated  
dagchainer.cpp:10:20: fatal error: stdio.h: No such file or directory  
 #include  <stdio.h>  
                    ^  
compilation terminated.  
make: *** [dagchainer.] Error 1  
``  

#ALTERNATE  
Ramos method:  
Identify all orthologous genes on scaffold = curated list of homologous genes (eg. 27 in Ramos)  
Classify as belonging to chromosome in comparison animal where gene of interest is found or somewhere else.  
If many paralogues, two tests: single location of each of the human orthologues, or collapsed location of human paralogues if all located at same locus  
Do a bunch of stuff to calculate probabilities.  
Do Exact binomial test and Fisher's test to test whether the number of correct neighbours is statistically significantly higher than expected by chance.    
**So....does order not matter?!!**    
-unless they wrote a script it seems very manual and time intensive.  

#Data/downloads
_Pleurobachia genome_:  https://www.ncbi.nlm.nih.gov/assembly/GCA_000695325.1  
P.bachei_draft_genome_v.1.1  
Organism: Pleurobrachia bachei (ctenophores)  
Submitter: University of Florida  
Date: 2014/05/21  
Assembly level: Scaffold  
Genome representation: full  
RefSeq category: representative genome  
GenBank assembly accession: GCA_000695325.1 (latest)  
RefSeq assembly accession: n/a  
IDs: 180401 [UID] 1073948 [GenBank]  
-paper said deposited at Moroz's website; links don't work on website.  

_Mnemiopsis leidyi_: https://www.ncbi.nlm.nih.gov/assembly/GCA_000226015.1/    
MneLei_Aug2011  
Organism name: Mnemiopsis leidyi (sea walnut)  
BioSample: SAMN02953801BioProject: PRJNA64405  
Submitter: National Human Genome Research Institute, National Institutes of Health  
Date: 2011/09/19  
Assembly level: Scaffold  
Genome representation: full  
RefSeq category: representative genome  
GenBank assembly accession: GCA_000226015.1 (latest)  
RefSeq assembly accession: n/a  
RefSeq assembly and GenBank assembly identical: n/a  
WGS Project: AGCP01  
Assembly method: Phusion v. 1.02  
Genome coverage: 12xSequencing technology: 454 GS-FLX Titanium; Illumina GA IIX  
IDs: 304208 [UID] 304208 [GenBank]  

_Amphimedon queenslandica: v1.0_: https://www.ncbi.nlm.nih.gov/assembly/GCF_000090795.1  
Organism: Amphimedon queenslandica (sponges)  
Submitter: US DOE Joint Genome Institute (JGI-PGF)   
Date: 2010/05/28  
Assembly level: Scaffold   
Genome representation: full  
RefSeq category: representative genome  
GenBank assembly accession: GCA_000090795.1 (latest)  
RefSeq assembly accession: GCF_000090795.1 (latest)  
IDs: 293608 [UID] 111438 [GenBank] 293608 [RefSeq]  

_Oscarella carmela_: http://www.compagen.org/datasets.html  OCAR not Oscarella sp.  
-added .fna and gzipped  
-Oscarella carmela (this assembly) renamed Oscarella pearsei; Oscarella sp. in Compgen redescribed Oscarella carmela. See 
https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0183002 - old O. carmela papers based on two species (carmela + pearsei). Both in compagen, but although there is  press release compagen didn't change their names in db. *Can't find annotation file*.  



#Papers+Links  
Riesgo et al., 2014  
https://academic.oup.com/mbe/article/31/5/1102/993377  
Neural genes Fig  
https://academic.oup.com/view-large/figure/74385341/msu057f3p.jpeg  

Moroz et al., 2014  
https://www.nature.com/articles/nature13400  
Table 34S: neural genes  
https://media.nature.com/original/nature-assets/nature/journal/v510/n7503/extref/nature13400-s1.pdf  

Ryan et al, 2013  
http://science.sciencemag.org/content/342/6164/1242592  
Suppl Mat:  
http://science.sciencemag.org/content/sci/suppl/2013/12/11/342.6164.1242592.DC1/Ryan.SM.pdf  

Alie and Manuel, 2010  
https://bmcevolbiol.biomedcentral.com/articles/10.1186/1471-2148-10-34  

Srivastava et al., 2010  
Suppl.S8.9 - neural genes
https://media.nature.com/original/nature-assets/nature/journal/v466/n7307/extref/nature09201-s1.pdf

Nichols et al., 2012  
https://www.pnas.org/content/109/32/13046  
Suppl  
https://www.pnas.org/content/pnas/suppl/2012/07/25/1120685109.DCSupplemental/sapp.pdf  




2. Should we be expecting these genes in these animals?  
Why should they use similar genes?  
How misguided is this approach? What is the true question implied by this approach?  


##References  
